---
title: "Validating automatic matching"
authors:
- affiliation: Department of Statistics and CSAFE, Iowa State University 
  name: Ganesh Krishnan
- affiliation: Department of Statistics and CSAFE, Iowa State University 
  name: Heike Hofmann
  thanks: The authors gratefully acknowledge funding from CSAFE through cooperative agreement \# 70NANB15H176 between NIST and Iowa State University. 
biblio-style: apsr
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
blinded: 0
keywords:
- at 
- least
- six
- keywords
bibliography: bibliography
abstract: 150 words abstract 
---


\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
\newcommand{\cited}[1]{{\textcolor{red}{#1}}}
\setlength\parindent{0pt}



\tableofcontents
\newpage
```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = TRUE,
  fig.path='figures/',
  echo=FALSE,
  cache=TRUE
)
options(knitr.table.format = "latex")
library(tidyverse)
library(kableExtra)
library(xtable)
library(gridExtra)
library(bulletr)
```

# Background and Motivation



- random forest \citep{aoas} trained on features extracted from land-to-land comparisons of Hamby 252 and Hamby 44 (now renamed to Hamby 173) available from NIST Ballistics Toolmarks Research Database (NBTRD) \citep{nist}
- scans taken on a NanoFocus  by Xiaoyu Zheng (Hamby 252) and Wei Chu (Hamby 44, now Hamby 173) at a resolution of 1.5625 $\mu m$. 

- Validation in several steps:
  - Hamby-44: different scanning method, same barrels
  - Phoenix PD: different scanning method, Ruger barrels
  - Houston: different scanning, Ruger, different model
  - Virginia: different ammunition, Smith & Wesson
  - LAPD: Beretta

- premise of the scores:
  - statistically meaningful
  - stable (i.e. the same score on one comparison means the same for another comparison)
  - explainable


```
nist_x3p %>% rotate_x3p(angle=-90) %>% y_flip_x3p() %>% image_x3p()
```

```{r}
library(bulletr)
hamby <- readr::read_csv("data/features-hamby.csv")
h44 <- readr::read_csv("data/h44-features.csv")
phoenix <- readr::read_csv("data/pd-features.csv")
houston <- readr::read_csv("data/houston-features.csv")

library(randomForest)
hamby$rfscore <- predict(rtrees, newdata = hamby, type="prob")[,2]
ones <- hamby %>% select(contains("1"))
twos <- hamby %>% select(contains("2"))
scores <- hamby %>% select(-c(names(ones), names(twos)))

hamby1 <- data.frame(ones, twos, scores)
hamby2 <- data.frame(twos, ones, scores)
names(hamby2) <- names(hamby1)

hamby <- rbind(hamby1, hamby2)

hamby %>% 
  filter(barrel1==1, barrel2 == barrel1, study1==study2, 
         study1=="Hamby252") %>% 
  ggplot(aes(x = land1, y = land2, fill=rfscore, colour=match)) +
  geom_tile(size=.5) +
  scale_fill_gradient2(low="grey30", midpoint=0.35, high="darkred") +
  scale_colour_manual(values=c(NA, "darkred")) +
  facet_grid(bullet1~bullet2) +
  theme(legend.position="bottom") +
  coord_equal()


h44$rfscore <- predict(rtrees, newdata = h44, type="prob")[,2]

bybullet <- h44 %>% 
  group_by(barrel1, barrel2, bullet1, bullet2) %>%
  nest()
bybullet <- bybullet %>% 
  mutate(
    bulletmatch = barrel1 == barrel2,
    bulletmatch = replace(bulletmatch, barrel1 == "Unk", NA),
    bulletmatch = replace(bulletmatch, barrel2 == "Unk", NA)
      )

bybullet <- bybullet %>% mutate(
  rfbullet = data %>% purrr::map_dbl(.f = function(d) {
    max(compute_average_scores(d$land1, d$land2, d$rfscore))
  }),
  rfindex = data %>% purrr::map_dbl(.f = function(d) {
    which.max(compute_average_scores(d$land1, d$land2, d$rfscore))
  })
)
h44 <- bybullet %>% unnest()
h44 <- h44 %>% mutate(
  match = ((land2-land1) %% 6 == rfindex-1) & (bullet1 != bullet2) & (bulletmatch | is.na(bulletmatch))
)

h44 <- h44 %>% mutate(
  potentialmatch = match,
  match = replace(match, is.na(bulletmatch), NA)
)

h44 %>% filter(barrel1 %in% c("1", "2"), barrel2 %in% c("1", "2")) %>% 
  ggplot(aes(x = land1, y = land2, fill=rfscore, colour = match)) +
  geom_tile(size=.5) +
  scale_fill_gradient2(low="grey30", midpoint=0.35, high="darkred") +
  scale_colour_manual(values=c(NA, "darkred")) +
  facet_grid(barrel1+bullet1~barrel2+bullet2, labeller="label_both") +
  theme(legend.position="bottom") +
  coord_equal()

h44 %>% filter(barrel1 %in% c("3", "4"), barrel2 %in% c("Unk")) %>% 
  ggplot(aes(x = land1, y = land2, fill=rfscore, colour = potentialmatch)) +
  geom_tile(size=.5) +
  scale_fill_gradient2(low="grey30", midpoint=0.35, high="darkred") +
  scale_colour_manual(values=c(NA, "darkred")) +
  facet_grid(barrel1+bullet1~barrel2+bullet2, labeller="label_both") +
  theme(legend.position="bottom") +
  coord_equal()

```



```{r, fig.cap="How similar are the extracted features between the different sets?"}
hamby %>% ggplot(aes(x = ccf)) + 
  geom_density(aes(colour= "NIST")) + 
  geom_density(data = h44, aes(colour= "Hamby-44")) +
  geom_density(data = phoenix, aes(colour= "Phoenix")) +
  geom_density(data = houston, aes(colour= "Houston")) 
```

# Appendix
\begin{appendix}




\end{appendix}